---
title:  The AusTraits Plant Dictionary (APD) Ontology
format: 
  html:
    embed-resources: true
    theme: cosmo
    smooth-scroll: true
    toc: true
    toc-expand: 1
---

# Release XX June, 2023

<dl>
<dt> **This version**:</dt> <dd><https://w3id.org/APD/2023-05-1/APD.html>  </dd>
<dt>**Latest version**:</dt> <dd><https://w3id.org/APD/APD.html> </dd>
<dt>**Latest version**:</dt> <dd><https://w3id.org/APD/APD.html> </dd>
<dt> **Previous version**:</dt> <dd> <https://w3id.org/APD/2023-02-1/APD.html> </dd>
<dt> **Authors**: </dt> <dd>[Elizabeth Wenk](https://orcid.org/0000-0001-5640-5910), [Daniel Falster](https://orcid.org/0000-0002-9814-092X"), [Hervé Sauquet](https://orcid.org/0000-0001-8305-3236), [Rowan Brownlee](https://orcid.org/0000-0002-1955-1262"), [Carl Boettiger](https://orcid.org/0000-0002-1642-628X") </dd>
<dt>**Contributors**:</dt> <dd> [Tony Auld](https://orcid.org/0000-0002-8766-2829),
  [Russell Barrett](https://orcid.org/0000-0003-0360-8321),
  [Timothy Brodribb](https://orcid.org/0000-0002-4964-6107),
  [Brendan Choat](https://orcid.org/0000-0002-9105-640X),
  [David Coleman](https://orcid.org/0000-0002-1773-6597),
  [Félix de Tombeur](https://orcid.org/0000-0002-6012-8458),
  [Lily Dun](https://orcid.org/0000-0002-0712-5143),
  [David Ellsworth](https://orcid.org/0000-0002-9699-2272),
  [Rachael Gallagher](https://orcid.org/0000-0002-4680-8115),
  [Carl Gosper](https://orcid.org/0000-0002-0962-5117),
  [Lydia Guja](https://orcid.org/0000-0001-5945-438X),
  [Gregory J. Jordan](https://orcid.org/0000-0002-6033-2766),
  [Tom Le Breton](https://orcid.org/0000-0001-9353-0067),
  [Andy Leigh](https://orcid.org/0000-0003-3568-2606),
  [Patricia Lu-Irving](https://orcid.org/0000-0003-1116-9402),
  [Belinda Medlyn](https://orcid.org/0000-0001-5728-9827),
  [Rachael Nolan](https://orcid.org/0000-0001-9277-5142),
  [Mark Ooi](https://orcid.org/0000-0002-3046-0417),
  [Karen D. Sommerville](https://orcid.org/0000-0002-9843-6616),
  [Ruby Stephens](https://orcid.org/0000-0002-3767-2690),
  [Peter Vesk](https://orcid.org/0000-0003-2008-7062),
  [Matthew White](https://orcid.org/0000-0003-2120-0071),
  [Ian J. Wright](https://orcid.org/0000-0001-8338-914),
  [Sophie Yang](https://orcid.org/0000-0001-7328-345X)</dd>

<dt>Download serialization:</dt> <dd>
<span><a href="ontology.json" target="_blank"><img src="https://img.shields.io/badge/Format-JSON_LD-blue.svg" alt="JSON-LD"></a> </span>
<span><a href="ontology.nt" target="_blank"><img src="https://img.shields.io/badge/Format-N_Triples-blue.svg" alt="N-Triples"></a> </span>
<span><a href="ontology.nq" target="_blank"><img src="https://img.shields.io/badge/Format-N_Quads-blue.svg" alt="N-Quads"></a> </span>
<span><a href="ontology.ttl" target="_blank"><img src="https://img.shields.io/badge/Format-TTL-blue.svg" alt="TTL"></a> </span></dd>

<dt>License: </dt> <dd><a href="http://purl.org/NET/rdflicense/cc-by4.0" target="_blank"><img src="https://img.shields.io/badge/License-CC%20BY%204.0-blue.svg" alt="http://purl.org/NET/rdflicense/cc-by4.0"></a> </dd>
<!-- <dt>Visualization:</dt><dd><a href="webvowl/index.html#" target="_blank"><img src="https://img.shields.io/badge/Visualize_with-WebVowl-blue.svg" alt="Visualize with WebVowl"></a></dd> -->

<dt>Cite as:</dt> <dd>Wenk et al, APD ontology, Retrieved from: https://w3id.org/APD/0.1.0</dd> </dl>

<!-- Examples:
https://i-adopt.github.io/index.html
https://www.w3.org/2009/08/skos-reference/skos.html
https://terminologies.gfbio.org/terms/ets/pages/#expectedunit

gt package: https://gt.rstudio.com/
-->


```{r, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(knitr)
library(kableExtra)
library(readr)
library(tidyr)
library(dplyr)
library(gt)
library(purrr)

source("R/kable.R")
source("R/helpers.R")
source("R/create_APD_trait_table.R")

triples_with_labels <- read_csv("data/triples_with_labels.csv")

if(knitr::is_html_output()) {
  my_table <- util_kable_styling_html
} else if (knitr::is_latex_output()){
  my_table <- util_kable_styling_pdf
} else {
  my_table <- util_kable_styling_docx
}


```
# Abstract
This document describes the schema available from the APD namespace. The APD offers detailed descriptions for more than 500 plant trait concepts. It includes trait focused on plant morphology, plant nutrient concentrations, plant physiology, plant life history, and plant fire response. The definitions will be useful to researchers from diverse disciplines, including plant functional ecology, plant taxonomy, and conservation biology. All trait concepts are supported by comprehensive metadata including trait descriptions, allowable trait values, allowable ranges, preferred units, keywords, references, and links to matches in a selection of trait databases. The traits describe here also fully support the AusTraits plant trait database, doi.org/10.5281/zenodo.3568417.

# Table of Contents 

1. [Introduction](#introduction)
2. [Trait groups](#trait-groupings): Hierarchical trait groupings, included to enhance trait findability.
3. [Trait concepts](#trait-concepts): Traits defined within the vocabulary.
4. [Values for categorical traits](#values-for-categorical-traits): Allowed categorical trait values to support the defined trait concepts.
5. [Glossary](#glossary): Terms used as keywords that were not available in other published vocabularies.
6. [Acknowledgements](#acknowledgements)

# 1. Introduction

The APD is a compilation of the traits concepts used by AusTraits, a database of Australian plant traits. APD includes definitions for nearly 500 traits pertaining to plant functional ecology and plant morphology. The metadata for each trait concept has been reviewed by multiple people. Metadata includes a trait definition, accepted units and ranges (for numeric traits), allowable values (for categorical traits), keywords and trait groupings to enhance trait findability, references, and links to identical/similar traits in other trait databases. APD will be published in a machine-readable format (rdf representation), allowing the traits to be readily re-used by other databases. This is a human-friendly output, with embedded links to the references, other trait databases and keywords/concepts borrowed from other ontologies.

The APD is a community resource and we welcome any suggestions for edits, including additional key references for each trait or other trait databases for which we should add links. 

# 2. Trait groups

The APD Trait groups express broad trait groupings, with each trait concept mapped into one or more trait groups as `narrower` concepts.

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
#trait_groupings <- datapasta::vector_paste(hierarchy_csv$Entity)

base_url <- "https://w3id.org/APD/traits/"

trait_groupings <- c("trait_0000000", "trait_0000002", "trait_0000004", "trait_0000006", "trait_0000008", "trait_0000108", "trait_0000206", "trait_0000408", "trait_0000508", "trait_0000608", "trait_0000808", "trait_0001006", "trait_0001502", "trait_0002004", "trait_0002005", "trait_0002006", "trait_0002007", "trait_0002008", "trait_0002009", "trait_0003004", "trait_0010001", "trait_0010004", "trait_0010114", "trait_0011001", "trait_0011202", "trait_0011204", "trait_0011206", "trait_0011304", "trait_0011402", "trait_0011502", "trait_0011503", "trait_0011602", "trait_0011702", "trait_0011802", "trait_0011902", "trait_0012002", "trait_0012006", "trait_0012102", "trait_0012104", "trait_0012106", "trait_0012108", "trait_0012202", "trait_0012404", "trait_0012406", "trait_0012408", "trait_0012409", "trait_0012504", "trait_0012604", "trait_0013001", "trait_0013006", "trait_0013102", "trait_0014001", "trait_0020001", "trait_0020202", "trait_0020204", "trait_0020206", "trait_0020208", "trait_0020210", "trait_0020212", "trait_0020812", "trait_0021002", "trait_0022002", "trait_0030001", "trait_0030006", "trait_0030008", "trait_0030009", "trait_0030059", "trait_0030504", "trait_0030602", "trait_0030702", "trait_0030704") %>% paste0(base_url, .)

to_match_tmp <- triples_with_labels %>%
  filter(Subject_stripped %in% trait_groupings) %>%
  filter(property == "label") %>%
  select(Subject_stripped, value)
 
for (thistrait in trait_groupings) {
  
  term_to_display <- to_match_tmp$value[match(thistrait,to_match_tmp$Subject_stripped)]
  
  ## [header1](#trait_0011212)

  anchor <- gsub(base_url, "", thistrait, fixed = TRUE)
  writeLines(paste0("\n## [", term_to_display, "](#",anchor,")\n\n"))

  out <- 
    create_APD_trait_hierarchy_table(thistrait, triples_with_labels)

  out %>% 
    gt() %>% 
    cols_align(align = "left") %>%
      cols_width(
        name ~ px(150),
        description ~px(500)
        ) %>%
      cols_label(
        name = "**Property**",
        description = "**Value**"
        ) %>%
    print()
}
```
# 3. Trait concepts

The APD Trait concepts express the traits defined by the APT.

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}

#datapasta::vector_paste(traits_csv$identifier)

base_url <- "https://w3id.org/APD/traits/"

list_of_traits <- c("trait_0000012", "trait_0000014", "trait_0000016", "trait_0000018", "trait_0000020", "trait_0000022", "trait_0000024", "trait_0000026", "trait_0000028", "trait_0000029", "trait_0000030", "trait_0000032", "trait_0000034", "trait_0000036", "trait_0000037", "trait_0000038", "trait_0000040", "trait_0000042", "trait_0000043", "trait_0000044", "trait_0000046", "trait_0000048", "trait_0000050", "trait_0000052", "trait_0000090", "trait_0000091", "trait_0000112", "trait_0000114", "trait_0000116", "trait_0000118", "trait_0000126", "trait_0000128", "trait_0000130", "trait_0000132", "trait_0000134", "trait_0000136", "trait_0000138", "trait_0000140", "trait_0000142", "trait_0000144", "trait_0000146", "trait_0000152", "trait_0022012", "trait_0022013", "trait_0000216", "trait_0000238", "trait_0000416", "trait_0000418", "trait_0000430", "trait_0000432", "trait_0000438", "trait_0000440", "trait_0000444", "trait_0000518", "trait_0000530", "trait_0000532", "trait_0000538", "trait_0000540", "trait_0000544", "trait_0000612", "trait_0000614", "trait_0000616", "trait_0000618", "trait_0000626", "trait_0000628", "trait_0000630", "trait_0000632", "trait_0000634", "trait_0000638", "trait_0000640", "trait_0000644", "trait_0000646", "trait_0000652", "trait_0000816", "trait_0000838", "trait_0000844", "trait_0001038", "trait_0001118", "trait_0001130", "trait_0001132", "trait_0001138", "trait_0001144", "trait_0001146", "trait_0001218", "trait_0001230", "trait_0001232", "trait_0001238", "trait_0001244", "trait_0001246",
                    "trait_0001511", "trait_0001512", "trait_0001513", "trait_0001514", "trait_0001611", "trait_0001612", "trait_0001613", "trait_0001614", "trait_0001615", "trait_0001616", "trait_0001661", "trait_0001662", "trait_0001663", "trait_0001664", "trait_0001665", "trait_0001666", "trait_0002021", "trait_0002022", "trait_0002024", "trait_0002025", "trait_0002027", "trait_0002028", "trait_0002031", "trait_0002032", "trait_0002035", "trait_0002037", "trait_0002050", "trait_0002052", "trait_0002054", "trait_0002055", "trait_0002056", "trait_0002081", "trait_0002082", "trait_0002083", "trait_0002084", "trait_0002085", "trait_0002086", "trait_0002087", "trait_0002090", "trait_0002127", "trait_0002131", "trait_0002224", "trait_0002250", "trait_0002255", "trait_0002327", "trait_0002331", "trait_0002534", "trait_0002544", "trait_0002822", "trait_0002824", "trait_0003011", "trait_0003012", "trait_0003013", "trait_0003014", "trait_0003015", "trait_0003031", "trait_0003032", "trait_0003033", "trait_0003034", "trait_0003035", "trait_0003052", "trait_0003053", "trait_0003072", "trait_0003092", "trait_0010021", "trait_0010022", "trait_0010023", "trait_0010024", "trait_0010025", "trait_0010070", "trait_0010110", "trait_0010111", "trait_0010112", "trait_0010113", "trait_0010114", "trait_0010160", "trait_0010161", "trait_0010162", "trait_0011211", "trait_0011212",
                    "trait_0011213", "trait_0011214", "trait_0011215", "trait_0011216", "trait_0011217", "trait_0011218", "trait_0011230", "trait_0011231", "trait_0011232", "trait_0011260", "trait_0011261", "trait_0011262", "trait_0011263", "trait_0011264", "trait_0011265", "trait_0011266", "trait_0011267", "trait_0011268", "trait_0011310", "trait_0011311", "trait_0011312", "trait_0011313", "trait_0011314", "trait_0011315", "trait_0011316", "trait_0011317", "trait_0011318", "trait_0011360", "trait_0011361", "trait_0011362", "trait_0011363", "trait_0011410", "trait_0011411", "trait_0011412", "trait_0011413", "trait_0011414", "trait_0011510", "trait_0011511", "trait_0011512", "trait_0011513", "trait_0011514", "trait_0011515", "trait_0011516", "trait_0011517", "trait_0011518", "trait_0011519", "trait_0011520", "trait_0011521", "trait_0011522", "trait_0011523", "trait_0011524", "trait_0011525", "trait_0011526", "trait_0011610", "trait_0011611", "trait_0011612", "trait_0011613", "trait_0011614", "trait_0011615", "trait_0011710", "trait_0011711", "trait_0011712", "trait_0011713", "trait_0011811", "trait_0011812", "trait_0011813", "trait_0011814", "trait_0011815", "trait_0011816", "trait_0011911", "trait_0011912",
                    "trait_0011913", "trait_0011914", "trait_0011915", "trait_0011916", "trait_0011917", "trait_0012010", "trait_0012011", "trait_0012012", "trait_0012013", "trait_0012014", "trait_0012015", "trait_0012016", "trait_0012111", "trait_0012112", "trait_0012113", "trait_0012114", "trait_0012115", "trait_0012116", "trait_0012117", "trait_0012118", "trait_0012119", "trait_0012120", "trait_0012121", "trait_0012122", "trait_0012221", "trait_0012222", "trait_0012410", "trait_0012411", "trait_0012412", "trait_0012413", "trait_0012414", "trait_0012415", "trait_0012416", "trait_0012417", "trait_0012418", "trait_0012419", "trait_0012420", "trait_0012421", "trait_0012422", "trait_0012431", "trait_0012432", "trait_0012433", "trait_0012434", "trait_0012441", "trait_0012442", "trait_0012443", "trait_0012444", "trait_0012445", "trait_0012461", "trait_0012462", "trait_0012463", "trait_0012464", "trait_0012465", "trait_0012466", "trait_0012467", "trait_0012471", "trait_0012472", "trait_0012473", "trait_0012481", "trait_0012482", "trait_0012483", "trait_0012484", "trait_0012511", "trait_0012512", "trait_0012513", "trait_0012514", "trait_0012515", "trait_0012516", "trait_0012517", "trait_0012518", "trait_0012519",
                    "trait_0012610", "trait_0012611", "trait_0012612", "trait_0012613", "trait_0012614", "trait_0012615", "trait_0012616", "trait_0012617", "trait_0012618", "trait_0012619", "trait_0012620", "trait_0012621", "trait_0012622", "trait_0012623", "trait_0012624", "trait_0013011", "trait_0013012", "trait_0013013", "trait_0013014", "trait_0013111", "trait_0013112", "trait_0013113", "trait_0013114", "trait_0013115", "trait_0013116", "trait_0013117", "trait_0013118", "trait_0013119", "trait_0013120", "trait_0013161", "trait_0013162", "trait_0013163", "trait_0013164", "trait_0013165", "trait_0014011", "trait_0014012", "trait_0014013", "trait_0014014", "trait_0014015", "trait_0014016", "trait_0014017", "trait_0014018", "trait_0014019", "trait_0014020", "trait_0020221", "trait_0020222", "trait_0020240", "trait_0020241", "trait_0020242", "trait_0020243", "trait_0020244", "trait_0020245", "trait_0020310", "trait_0020311", "trait_0020312", "trait_0020313", "trait_0020314", "trait_0020315", "trait_0020410", "trait_0020411", "trait_0020412", "trait_0020413", "trait_0020414", "trait_0020415", "trait_0020416", "trait_0020417", "trait_0020418", "trait_0020419", "trait_0020510", "trait_0020511", "trait_0020512",
                    "trait_0020513", "trait_0020610", "trait_0020611", "trait_0020612", "trait_0020630", "trait_0020640", "trait_0020641", "trait_0020660", "trait_0020661", "trait_0020662", "trait_0020663", "trait_0020664", "trait_0020710", "trait_0020711", "trait_0020712", "trait_0020713", "trait_0020760", "trait_0020761", "trait_0020762", "trait_0021013", "trait_0021014", "trait_0021015", "trait_0021016", "trait_0021021", "trait_0021022", "trait_0021030", "trait_0021031", "trait_0021040", "trait_0021050", "trait_0021051", "trait_0021052", "trait_0021053", "trait_0021054", "trait_0021055", "trait_0021056", "trait_0021057", "trait_0021058", "trait_0021059", "trait_0020816", "trait_0020813", "trait_0020814", "trait_0020815", "trait_0030010", "trait_0030011", "trait_0030012", "trait_0030013", "trait_0030014", "trait_0030015", "trait_0030016", "trait_0030017", "trait_0030018", "trait_0030019", "trait_0030020", "trait_0030021", "trait_0030022", "trait_0030023", "trait_0030024", "trait_0030025", "trait_0030026", "trait_0030027", "trait_0030028", "trait_0030060", "trait_0030061", "trait_0030110", "trait_0030111", "trait_0030260", "trait_0030261", "trait_0030310", "trait_0030311", "trait_0030210", "trait_0030211", 
                    "trait_0030313", "trait_0030314", "trait_0030315", "trait_0030410", "trait_0030511", "trait_0030512", "trait_0030513", "trait_0030514", "trait_0030515", "trait_0030516", "trait_0030517", "trait_0030530", "trait_0030531", "trait_0030532", "trait_0030533", "trait_0030560", "trait_0030561", "trait_0030590", "trait_0030591", "trait_0030610", "trait_0030612", "trait_0030613", "trait_0030614", "trait_0030615", "trait_0030616", "trait_0030621", "trait_0030622", "trait_0030623", "trait_0030624", "trait_0030625", "trait_0030626", "trait_0030627", "trait_0030631", "trait_0030632", "trait_0030633", "trait_0030634", "trait_0030635", "trait_0030636", "trait_0030637", "trait_0030710", "trait_0030711", "trait_0030712", "trait_0030713", "trait_0030714", "trait_0030715", "trait_0030716", "trait_0030717", "trait_0030718", "trait_0030619") %>%
  paste0(base_url, .)

to_match_tmp <- triples_with_labels %>%
  filter(Subject_stripped %in% list_of_traits) %>%
  filter(property == "label") %>%
  select(Subject_stripped, value)
 
for (thistrait in list_of_traits) {
  
  term_to_display <- to_match_tmp$value[match(thistrait,to_match_tmp$Subject_stripped)]
  
  ## [header1](#trait_0011212)

  anchor <- gsub(base_url, "", thistrait, fixed = TRUE)
  writeLines(paste0("\n## [", term_to_display, "](#",anchor,")\n\n"))

  out <- 
    create_APD_trait_table(thistrait, triples_with_labels)

  out %>% 
    gt() %>% 
    cols_align(align = "left") %>%
    cols_width(
        name ~ px(150),
        description ~px(600)
      ) %>%
    cols_label(
        name = "**Property**",
        description = "**Value**"
        ) %>%
    print()
}
```


# 4. Values for categorical traits

This section of the ontology expresses allowed categorical trait values for each trait concept. These terms are each defined and linked to a single trait concept to reduce ambiguity.

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
#datapasta::vector_paste(categorical_values_csv$identifier)

base_url <- "https://w3id.org/APD/traits/"

categorical_allowed <- c("bark_morphology_eucalyptus_eucalypt_box", "bark_morphology_eucalyptus_eucalypt_gum", "bark_morphology_eucalyptus_eucalypt_stocking", "bark_morphology_eucalyptus_eucalypt_ironbark", "bark_morphology_eucalyptus_eucalypt_peppermint", "bark_morphology_eucalyptus_eucalypt_stringybark", "bark_morphology_eucalyptus_eucalypt_ribbonbark", "plant_tolerance_calcicole_calcicole", "plant_tolerance_calcicole_calcifuge", "competitive_stratum_ground", "competitive_stratum_mid", "competitive_stratum_upper", "competitive_stratum_post_fire_ephemeral", "cotyledon_function_foliaceous", "cotyledon_function_reserve", "cotyledon_position_cryptocotylar", "cotyledon_position_phanerocotylar", "cotyledon_hairs_glabrous", "cotyledon_hairs_hairy", "seedling_germination_location_epigeal", "seedling_germination_location_hypogeal", "dispersal_appendage_aril", "dispersal_appendage_bristles", "dispersal_appendage_bracts_and_glumes", "dispersal_appendage_elaiosome", "dispersal_appendage_fleshy_reward", "dispersal_appendage_floating_seed", "dispersal_appendage_flotation_scales", "dispersal_appendage_floral_parts", "dispersal_appendage_hairs", "dispersal_appendage_hooks", "dispersal_appendage_inflated_parts", "dispersal_appendage_none", "dispersal_appendage_pappus", "dispersal_appendage_plumose", "dispersal_appendage_receptacle", "dispersal_appendage_sarcotesta", "dispersal_appendage_spines", "dispersal_appendage_tumbleweed", "dispersal_appendage_wings", "dispersal_appendage_wing_or_plume", "dispersal_syndrome_anemochory", "dispersal_syndrome_anthrochory", "dispersal_syndrome_atelochory", "dispersal_syndrome_autochory", "dispersal_syndrome_ballistic",
"dispersal_syndrome_barochory", "dispersal_syndrome_chamaechory", "dispersal_syndrome_endozoochory", "dispersal_syndrome_epizoochory", "dispersal_syndrome_hydrochory", "dispersal_syndrome_myrmecochory", "dispersal_syndrome_undefined", "dispersal_syndrome_zoochory", "dispersal_unit_fruit", "dispersal_unit_seed", "dispersal_unit_spore", "dispersers_indigenous_people", "dispersers_ants", "dispersers_bats", "dispersers_birds", "dispersers_cassowaries", "dispersers_floods", "dispersers_flying_vertebrates", "dispersers_flying_foxes", "dispersers_garden_refuse", "dispersers_insects", "dispersers_invertebrates", "dispersers_mammals", "dispersers_non-flying_vertebrates", "dispersers_passive", "dispersers_rain", "dispersers_rodents", "dispersers_vehicles", "dispersers_vertebrates", "dispersers_wind", "dispersers_water", "dispersers_water_currents", "seed_dormancy_class_morphophysiological_dormancy", "seed_dormancy_class_non_dormant", "seed_dormancy_class_physical_dormancy", "seed_dormancy_class_physiological_dormancy", "embryo_colour_colourless", "embryo_colour_green", "fire_exposure_level_fire_not_relevant", "fire_exposure_level_fire_avoidance_among_rocks", "resprouting_capacity_fire_killed", "resprouting_capacity_resprouts", "resprouting_capacity_partial_resprouting", "resprouting_capacity_no_significant_resprouting_capacity", "resprouting_capacity_juvenile_juvenile_fire_killed", "resprouting_capacity_juvenile_juvenile_resprout", "resprouting_capacity_juvenile_juvenile_moderate_fire_killed",
"fire_response_on_maturity_remain_mature", "fire_response_on_maturity_mature_to_juvenile", "fire_response_on_maturity_mature_do_not_resprout", "post_fire_recruitment_post_fire_recruitment", "post_fire_recruitment_post_fire_recruitment_moderate", "post_fire_recruitment_no_post_fire_recruitment", "flower_colour_blue_purple", "flower_colour_green", "flower_colour_pink", "flower_colour_red_brown", "flower_colour_white_cream", "flower_colour_yellow_orange", "flowering_cues_rain", "flowering_cues_rain_obligate", "flowering_cues_rain_all_year", "flowering_cues_rain_summer", "flowering_cues_rain_autumn", "flowering_cues_rain_winter", "flowering_cues_rain_spring", "flowering_cues_floods", "flowering_cues_fire", "flower_structural_sex_type_bisexual", "flower_structural_sex_type_incompletely_unisexual", "flower_structural_sex_type_unisexual", "flower_ovary_position_inferior", "flower_ovary_position_superior", "flower_ovary_position_half_inferior", "flower_ovary_position_one_quarter_inferior", "flower_ovary_position_three_quarters_inferior", "flower_perianth_symmetry_actinomorphic_general", "flower_perianth_symmetry_actinomorphic_rotational", "flower_perianth_symmetry_actinomorphic_strictly_polysymmetric", "flower_perianth_symmetry_actinomorphic_spiral", "flower_perianth_symmetry_asymmetric", "flower_perianth_symmetry_disymmetric", "flower_perianth_symmetry_zygomorphic", "flower_perianth_phyllotaxis_irregular", "flower_perianth_phyllotaxis_spiral", "flower_perianth_phyllotaxis_whorled", "flower_perianth_differentiation_undifferentiated", "flower_perianth_differentiation_marked_differentiation_outer_sepaloid_inner_petaloid", "flower_perianth_differentiation_marked_differentiation_among_petaloid_tepals",
"flower_perianth_differentiation_marked_differentiation_among_sepaloid_tepals", "flower_perianth_differentiation_marked_differentiation", "flower_perianth_differentiation_continuous_differentiation_outer_sepaloid_inner_petaloid", "flower_perianth_differentiation_continuous_differentiation_among_petaloid_tepals", "flower_perianth_differentiation_continuous_differentiation_among_sepaloid_tepals", "flower_perianth_differentiation_weak_differentiation", "flower_perianth_differentiation_very_weak_differentiation", "flower_filament_absent", "flower_filament_present_general", "flower_filament_long_and_narrow", "flower_filament_short_and_narrow", "flower_filament_long_and_wide", "flower_filament_short_and_wide", "flower_filament_long_and_very_wide", "flower_filament_petaloid", "flower_filament_fused_into_synandrium", "flower_filament_fused_into_fascicle_stalks", "flower_filament_long_and_narrow_divided_in_two", "flower_filament_short_general", "flower_anther_orientation_apical", "flower_anther_orientation_extrorse", "flower_anther_orientation_introrse", "flower_anther_orientation_latrorse", "flower_anther_attachment_basifixed", "flower_anther_attachment_dorsifixed", "flower_anther_attachment_dorsifixed_at_base", "flower_anther_attachment_ventrifixed", "flower_anther_dehiscence_longitudinal_slit", "flower_anther_dehiscence_poricidal", "flower_anther_dehiscence_h_valvate", "flower_anther_dehiscence_flap_valvate", "flower_anther_dehiscence_transverse_slit", "flower_anther_dehiscence_short_basal_slits", "flower_anther_dehiscence_short_central_slits", "flower_anther_dehiscence_short_apical_slits", "flower_anther_dehiscence_t_valvate", "flower_anther_dehiscence_common_stomium_confluent_thecae", "flower_anther_connective_extension_absent", "flower_anther_connective_extension_short_extension", "flower_anther_connective_extension_long_extension", "flower_anther_connective_extension_very_long_extension", "flower_anther_connective_extension_present_general", "flower_androecium_structural_phyllotaxis_whorled", "flower_androecium_structural_phyllotaxis_spiral", "flower_androecium_structural_phyllotaxis_irregular", "flower_gynoecium_phyllotaxis_one_whorl", "flower_gynoecium_phyllotaxis_two_or_more_whorls", "flower_gynoecium_phyllotaxis_spiral", "flower_style_differentiation_absent",
"flower_style_differentiation_present_length_and_shape_unknown", "flower_style_differentiation_present_short_and_thick", "flower_style_differentiation_present_short_and_narrow", "flower_style_differentiation_present_long_and_wide", "flower_style_differentiation_present_long_and_narrow", "flower_style_differentiation_present_petaloid", "flower_style_differentiation_continuous", "flower_gynoecium_placentation_axile", "flower_gynoecium_placentation_apical", "flower_gynoecium_placentation_basal", "flower_gynoecium_placentation_parietal", "flower_gynoecium_placentation_marginal", "flower_gynoecium_placentation_free_central", "flower_gynoecium_placentation_laminar", "flower_pollen_aperture_shape_absent", "flower_pollen_aperture_shape_colporus", "flower_pollen_aperture_shape_colpus", "flower_pollen_aperture_shape_pore", "flower_pollen_aperture_shape_ring_like", "flower_pollen_aperture_shape_spiraperturate", "flower_pollen_aperture_shape_sulcus", "flower_pollen_aperture_shape_syncolpus", "flower_pollen_aperture_shape_ulcus", "fruit_colour_pink", "fruit_colour_red", "fruit_colour_orange", "fruit_colour_yellow", "fruit_colour_green", "fruit_colour_blue", "fruit_colour_purple", "fruit_colour_white", "fruit_colour_cream", "fruit_colour_brown", "fruit_colour_black", "fruit_colour_grey", "fruit_type_achene", "fruit_type_anthocarp", "fruit_type_berry", "fruit_type_capsule", "fruit_type_caryopsis", "fruit_type_compound_fruit_other", "fruit_type_drupe", "fruit_type_follicle", "fruit_type_legume", "fruit_type_legume_indehiscent", "fruit_type_mericarp", "fruit_type_multiple_fruit", "fruit_type_nut", "fruit_type_nutlet", "fruit_type_pepo", "fruit_type_pome", "fruit_type_pyrene",
"fruit_type_samara", "fruit_type_schizocarp", "fruit_type_silique", "fruit_type_strobilus", "fruit_type_syncarp", "fruit_type_syconium", "fruit_type_utricle", "fruit_fleshiness_dry", "fruit_fleshiness_fleshy", "fruit_dehiscence_dehiscent", "fruit_dehiscence_indehiscent", "diaspore_fleshiness_dry", "diaspore_fleshiness_fleshy", "seed_germination_treatment_after_ripening", "seed_germination_treatment_chemical_treatment", "seed_germination_treatment_cold", "seed_germination_treatment_dark", "seed_germination_treatment_heat", "seed_germination_treatment_heat_lethal", "seed_germination_treatment_heat+smoke", "seed_germination_treatment_imbibe_hot", "seed_germination_treatment_leaching", "seed_germination_treatment_light", "seed_germination_treatment_mycorrhizae", "seed_germination_treatment_none", "seed_germination_treatment_rain", "seed_germination_treatment_scarify", "seed_germination_treatment_smoke", "seed_germination_treatment_stratification", "seed_germination_treatment_stratification_cold", "seed_germination_treatment_stratification_warm", "seedling_hypocotyl_hairs_glabrous", "seedling_hypocotyl_hairs_hairy", "plant_tolerance_inundation_not_applicable", "plant_tolerance_inundation_less_than_1_month", "plant_tolerance_inundation_1-6_months", "plant_tolerance_inundation_greater_than_6_months", "plant_tolerance_inundation_aquatic", "plant_flood_regime_classification_terrestrial_dry", "plant_flood_regime_classification_terrestrial_damp", "plant_flood_regime_classification_amphibious_fluctuation_tolerators_emergent", "plant_flood_regime_classification_amphibious_fluctuation_tolerators_low", "plant_flood_regime_classification_amphibious_fluctuation_responders_plastic",
"plant_flood_regime_classification_amphibious_fluctuation_responders_floating", "plant_flood_regime_classification_amphibious", "plant_flood_regime_classification_submerged", "leaf_pendulousness_pendulous", "leaf_pendulousness_not_pendulous", "leaf_arrangement_decussate", "leaf_arrangement_distichous", "leaf_arrangement_spiral", "leaf_arrangement_fasciculate", "leaf_arrangement_crowded", "leaf_phyllotaxis_alternate", "leaf_phyllotaxis_opposite", "leaf_phyllotaxis_whorled", "leaf_glaucousness_glaucous", "leaf_glaucousness_not_glaucous", "leaf_hairs_adult_leaves_glabrous", "leaf_hairs_adult_leaves_hairy", "leaf_hairs_immature_leaves_glabrous", "leaf_hairs_immature_leaves_hairy", "leaf_hairs_juvenile_leaves_glabrous", "leaf_hairs_juvenile_leaves_hairy", "leaf_margin_entire", "leaf_margin_toothed", "leaf_margin_toothed_serrate", "leaf_margin_toothed_dentate", "leaf_margin_toothed_crenate", "leaf_margin_posture_revolute", "leaf_margin_posture_recurved", "leaf_margin_posture_flat", "leaf_margin_posture_undulate", "leaf_lobation_unlobed", "leaf_lobation_lobed", "leaf_lobation_lobed_deep", "leaf_lobation_lobed_shallow", "leaf_compoundness_simple", "leaf_compoundness_compound", "leaf_lamina_division_pinnately_lobed", "leaf_lamina_division_palmately_lobed", "leaf_lamina_division_pinnatifid", "leaf_lamina_division_bipinnatifid", "leaf_lamina_division_pinnatipartite", "leaf_lamina_division_pinnatisect", "leaf_lamina_division_bipinnatipartite", "leaf_lamina_division_bipinnatisect", "leaf_lamina_division_trichotomously_lobed", "leaf_lamina_division_dichotomously_lobed",
"leaf_lamina_division_bipartite", "leaf_lamina_division_tripartite", "leaf_lamina_division_pinnately_compound", "leaf_lamina_division_palmately_compound", "leaf_lamina_division_bipinnate", "leaf_lamina_division_tripinnate", "leaf_lamina_division_trifoliate", "leaf_phenology_deciduous", "leaf_phenology_evergreen", "leaf_phenology_cold_deciduous", "leaf_phenology_drought_deciduous", "leaf_phenology_facultative_drought_deciduous", "leaf_phenology_semi_deciduous", "leaf_phenology_brevi_deciduous", "leaf_shape_linear", "leaf_shape_narrowly_oblong", "leaf_shape_oblong", "leaf_shape_linear_elliptical", "leaf_shape_narrowly_elliptical", "leaf_shape_elliptical", "leaf_shape_widely_elliptical", "leaf_shape_orbicular", "leaf_shape_oblate", "leaf_shape_lanceolate", "leaf_shape_narrowly_ovate", "leaf_shape_ovate", "leaf_shape_widely_ovate", "leaf_shape_narrowly_oblanceolate", "leaf_shape_oblanceolate", "leaf_shape_narrowly_obovate", "leaf_shape_obovate", "leaf_shape_widely_obovate", "leaf_shape_narrowly_triangular", "leaf_shape_triangular", "leaf_shape_deltate", "leaf_shape_obtriangular", "leaf_shape_narrowly_rhomboidal", "leaf_shape_rhomboidal", "leaf_shape_trullate", "leaf_shape_narrowly_obtrullate", "leaf_shape_obtrullate", "leaf_shape_widely_obtrullate", "leaf_shape_acicular", "leaf_shape_terete", "leaf_shape_cordate", "leaf_shape_obcordate", "leaf_shape_falcate", "leaf_shape_reniform",
"leaf_shape_spathulate", "leaf_shape_subulate", "leaf_shape_peltate", "leaf_shape_filiform", "leaf_base_shape_attenuate", "leaf_base_shape_auriculate", "leaf_base_shape_cuneate", "leaf_base_shape_hastate", "leaf_base_shape_sagittate", "plant_photosynthetic_organ_leaf", "plant_photosynthetic_organ_cladode", "plant_photosynthetic_organ_phyllode", "plant_photosynthetic_organ_non-photosynthetic_plant", "leaf_type_broadleaf", "leaf_type_needle", "leaf_type_scale", "leaf_type_leafless", "life_form_phanerophyte", "life_form_chamaephyte", "life_form_hemicryptophyte", "life_form_cryptophyte", "life_form_geophyte", "life_form_helophyte", "life_form_hydrophyte", "life_form_therophyte", "life_form_epiphyte", "life_form_aerophyte", "life_history_ephemeral", "life_history_annual", "life_history_biennial", "life_history_perennial", "life_history_short_lived_perennial", "sex_type_androdioecious", "sex_type_andromonoecious", "sex_type_dioecious", "sex_type_gynodioecious", "sex_type_gynomonoecious", "sex_type_hermaphrodite", "sex_type_monoecious", "sex_type_polygamous", "sex_type_polygamodioecious", "sex_type_polygamonoecious", "sex_type_trioecious", "sex_type_unknown", "nitrogen_fixing_non_nitrogen_fixer", "nitrogen_fixing_nitrogen_fixer", "parasitic_hemiparasitic", "parasitic_hemiparasitic_obligate", "parasitic_hemiparasitic_facultative", "parasitic_holoparasitic", "parasitic_parasitic", "parasitic_not_parasitic", "parasitic_root_parasitic", "parasitic_stem_parasitic", "plant_alternative_energy_and_nutrient_acquisition_strategy_saprophyte", "bark_photosynthetic_status_non_photosynthetic_bark", "bark_photosynthetic_status_photosynthetic_bark", "photosynthetic_pathway_c3", "photosynthetic_pathway_c4", "photosynthetic_pathway_c3-c4", "photosynthetic_pathway_cam", "photosynthetic_pathway_c3-cam", "photosynthetic_pathway_c4-cam", "photosynthetic_pathway_facultative_cam", "photosynthetic_pathway_unknown", "plant_growth_form_herb", "plant_growth_form_subshrub", "plant_growth_form_shrub", "plant_growth_form_mallee", "plant_growth_form_tree", "plant_growth_form_graminoid", "plant_growth_form_tussock",
"plant_growth_form_graminoid_not_tussock", "plant_growth_form_hummock", "plant_growth_form_basal_large", "plant_growth_form_geophyte", "plant_growth_form_palmoid", "plant_growth_form_climber", "plant_growth_form_climber_herbaceous", "plant_growth_form_climber_woody", "plant_growth_form_fern", "plant_growth_form_lycophyte", "plant_succulence_succulent_stems", "plant_succulence_succulent_leaves", "plant_succulence_succulent", "plant_succulence_semi-succulent", "plant_succulence_not_succulent", "stem_growth_habit_arborescent", "stem_growth_habit_erect", "stem_growth_habit_caespitose", "stem_growth_habit_cushion-forming", "stem_growth_habit_mat-forming", "stem_growth_habit_tufted", "stem_growth_habit_creeping", "stem_growth_habit_decumbent", "stem_growth_habit_prostrate", "stem_growth_habit_rosette", "stem_growth_habit_climbing", "stem_growth_habit_sprawling", "stem_growth_habit_rhizomatous", "stem_growth_habit_stoloniferous", "stem_growth_habit_spreading", "stem_growth_habit_floating", "stem_growth_habit_submerged", "stem_growth_habit_dense", "stem_growth_habit_compact", "stem_growth_habit_bushy", "stem_growth_habit_open", "stem_growth_habit_spindly", "plant_climbing_mechanism_adventitious_roots", "plant_climbing_mechanism_hooks", "plant_climbing_mechanism_spines", "plant_climbing_mechanism_tendrils", "plant_climbing_mechanism_scrambling", "plant_climbing_mechanism_twining", "plant_growth_substrate_epiphyte", "plant_growth_substrate_lithophyte", "plant_growth_substrate_terrestrial", "plant_growth_substrate_semiaquatic", "plant_growth_substrate_aquatic", "plant_growth_substrate_marine", "plant_growth_substrate_hemiepiphyte", "plant_type_by_resource_use_hydrohalophyte", "plant_type_by_resource_use_phreatophyte", "plant_type_by_resource_use_psammophile", "plant_type_by_resource_use_weedy", "plant_type_by_resource_use_xerohalophyte", "plant_type_by_resource_use_xerophyte", "pollination_syndrome_abiotic", "pollination_syndrome_biotic", "pollination_syndrome_animal", "pollination_syndrome_bat", "pollination_syndrome_bee", "pollination_syndrome_beetle", "pollination_syndrome_bird", "pollination_syndrome_butterfly", "pollination_syndrome_fly", "pollination_syndrome_insect", "pollination_syndrome_mammal", "pollination_syndrome_moth", "pollination_syndrome_thrip", "pollination_syndrome_vertebrate", "pollination_syndrome_wasp", "pollination_syndrome_water", "pollination_syndrome_wind", "pollination_syndrome_self",
"pollination_syndrome_pollination_rare", "pollination_system_abiotic", "pollination_system_biotic_unspecialised", "pollination_system_biotic_specialised", "pollination_system_animal", "pollination_system_bat", "pollination_system_bee", "pollination_system_beetle", "pollination_system_bird", "pollination_system_butterfly", "pollination_system_fly", "pollination_system_insect", "pollination_system_mammal", "pollination_system_moth", "pollination_system_thrip", "pollination_system_vertebrate", "pollination_system_wasp", "pollination_system_water", "pollination_system_wind", "pollination_system_self", "regeneration_non_fire_disturbance_resprouts_non_fire_disturbance", "root_system_classification_type_i", "root_system_classification_type_ii", "root_system_classification_type_iii", "root_system_classification_type_iv", "root_system_classification_type_v", "root_system_classification_type_vi", "root_system_classification_type_vii", "root_system_classification_type_viii", "root_system_classification_type_ix", "root_structure_arbuscular_mycorrhizal", "root_structure_carnivorous", "root_structure_cluster_roots", "root_structure_coiling_vam", "root_structure_dauciform_root", "root_structure_ectomycorrhizal", "root_structure_ericoid_mycorrhizal", "root_structure_fine_roots", "root_structure_haustorium", "root_structure_hemiparasitic_root", "root_structure_long_root_hairs", "root_structure_mycorrhizal", "root_structure_non_mycorrhizal", "root_structure_orchid_mycorrhizal", "root_structure_parasitic_root", "root_structure_proteoid_root", "root_structure_root_hairs", "root_structure_sand-binding", "root_structure_saprophyte", "root_structure_subepidermal_mycorrhizal", "root_structure_unknown", "plant_tolerance_salt_glycophyte", "plant_tolerance_salt_halophyte", "plant_tolerance_salt_halophyte_moderate", "plant_tolerance_salt_hydrohalophyte", "plant_tolerance_salt_salt_spray_tolerant",
"plant_tolerance_salt_salinity_tolerance_undefined", "plant_tolerance_salt_xerohalophyte", "seed_shape_comma-shaped", "seed_shape_conical", "seed_shape_cuneate", "seed_shape_cylindrical", "seed_shape_discoid", "seed_shape_ellipsoid", "seed_shape_ellipsoid_flattened", "seed_shape_fusiform", "seed_shape_globoid", "seed_shape_globoid_flattened", "seed_shape_hemispheric", "seed_shape_hemispheric_flattened", "seed_shape_lenticular", "seed_shape_orbicular", "seed_shape_ovoid", "seed_shape_ovoid_elongated", "seed_shape_ovoid_flattened", "seed_shape_polyhedral", "seed_shape_polyhedral_elongated", "seed_shape_polyhedral_inflated", "seed_shape_reniform", "seed_shape_winged", "seed_shape_sectoroid", "seed_shape_sinusoidal", "seed_surface_hairs_glabrous", "seed_surface_hairs_hairs", "seed_surface_hairs_hairs_branched", "seed_surface_hairs_hairs_simple", "seed_surface_reflectivity_dull", "seed_surface_reflectivity_shiny", "seed_surface_texture_bumpy", "seed_surface_texture_netted", "seed_surface_texture_papery", "seed_surface_texture_pitted", "seed_surface_texture_ribbed", "seed_surface_texture_rough", "seed_surface_texture_scaly", "seed_surface_texture_smooth", "seed_surface_texture_spiny", "seed_surface_texture_wrinkled", "canopy_seedbank_canopy_seedbank_absent", "canopy_seedbank_canopy_seedbank_present", "soil_seedbank_soil_seedbank_absent", "soil_seedbank_soil_seedbank_present", "seed_storage_location_canopy", "seed_storage_location_soil", "seed_storage_location_none", "seed_storage_location_not_canopy", "seed_longevity_categorical_transient",
"seed_longevity_categorical_intermediate_longevity", "seed_longevity_categorical_persistent", "seed_longevity_categorical_short_persistent", "seed_longevity_categorical_long_persistent", "seed_longevity_categorical_widely_dispersed", "seed_release_serotinous", "seed_release_mature_en_mass", "seed_release_gradual_release", "serotiny_not_serotinous", "serotiny_serotinous", "serotiny_serotiny_low", "serotiny_serotiny_moderate", "serotiny_serotiny_high", "seedling_first_node_leaf_type_scale-like", "seedling_first_node_leaf_type_leaf-like", "seedling_first_node_leaf_count_single", "seedling_first_node_leaf_count_paired", "plant_tolerance_snow_not_applicable", "plant_tolerance_snow_intolerant", "plant_tolerance_snow_days", "plant_tolerance_snow_weeks", "plant_tolerance_snow_months", "plant_physical_defence_structures_absent", "plant_physical_defence_structures_sharp_pointed_defence", "plant_physical_defence_structures_prickle", "plant_physical_defence_structures_pungent_leaf_apex", "plant_physical_defence_structures_spine",
"plant_physical_defence_structures_stinging_or_irritant_hairs", "plant_physical_defence_structures_thorn", "plant_spinescence_no_spines", "plant_spinescence_low_density_soft_spines", "plant_spinescence_high_density_soft_spines", "plant_spinescence_high_density_hard_spines_to_5mm", "plant_spinescence_high_density_hard_spines_to_20mm", "plant_spinescence_high_density_hard_spines_to_100mm", "leaf_stomatal_distribution_amphistomatic", "leaf_stomatal_distribution_amphistomatic_uneven", "leaf_stomatal_distribution_epistomatic", "leaf_stomatal_distribution_hypostomatic", "leaf_stomatal_hairs_stomata_covered_with_dense_hairs", "leaf_stomatal_hairs_stomata_not_covered_by_hairs", "root_system_type_tap_root_absent", "root_system_type_tap_root_present", "root_system_type_tap_root_sometimes_present", "vegetative_reproduction_ability_not_vegetative", "vegetative_reproduction_ability_vegetative", "plant_tolerance_water_logged_soils_not_applicable", "plant_tolerance_water_logged_soils_less_than_1_month", "plant_tolerance_water_logged_soils_1-6_months", "plant_tolerance_water_logged_soils_greater_than_6_months", "plant_tolerance_water_logged_soils_aquatic", "woodiness_herbaceous", "woodiness_woody", "woodiness_detailed_herbaceous", "woodiness_detailed_woody", "woodiness_detailed_woody_base", "woodiness_detailed_woody_root", "woodiness_detailed_semi_woody", "woodiness_detailed_woody_like_stem", "woodiness_detailed_woody_like_inflorescence", "bud_bank_location_apical_buds", "bud_bank_location_basal_burl", "bud_bank_location_basal_stem_buds", "bud_bank_location_epicormic_buds", "bud_bank_location_fleshy_underground_organ", "bud_bank_location_root_crown", "bud_bank_location_bud-bearing_root", "bud_bank_location_rhizome", "bud_bank_location_rhizome_woody", "bud_bank_location_rhizome_non_woody", "bud_bank_location_stem_buds", "bud_bank_location_stem_tuber", "bud_bank_location_stolon", "bud_bank_location_underground_buds", "bud_bank_location_none", "clonal_spread_mechanism_aboveground_clonal",
"clonal_spread_mechanism_adventitious_roots", "clonal_spread_mechanism_bulb", "clonal_spread_mechanism_belowground_clonal", "clonal_spread_mechanism_corm", "clonal_spread_mechanism_rhizome", "clonal_spread_mechanism_root_buds", "clonal_spread_mechanism_root_tuber", "clonal_spread_mechanism_stem_tuber", "clonal_spread_mechanism_stolon", "clonal_spread_mechanism_viviparous", "clonal_spread_mechanism_non_clonal", "storage_organ_bulb", "storage_organ_corm", "storage_organ_caudex", "storage_organ_fleshy_underground_organ", "storage_organ_lignotuber", "storage_organ_rhizome", "storage_organ_rhizome_woody", "storage_organ_rhizome_felshy", "storage_organ_rhizome_fibrous", "storage_organ_root_tuber", "storage_organ_stem_tuber", "storage_organ_storage_organ", "storage_organ_no_storage_organ", "life_history_ephemeral_class_disturbance_ephemeral", "life_history_ephemeral_class_rain_ephemeral", "life_history_ephemeral_class_fire_ephemeral", "life_history_ephemeral_class_fire_ephemeral_facultative", "life_history_ephemeral_class_fire_ephemeral_obligate", "post_fire_flowering_fire_dependent_flowering", "post_fire_flowering_fire_enhanced_flowering",
"post_fire_flowering_fire_independent_flowering", "post_fire_flowering_fire_suppressed_flowering", "plant_influence_on_fire_risk_fire_risk_decreased", "plant_influence_on_fire_risk_fire_risk_increased", "plant_influence_on_fire_risk_fire_risk_unchanged", "plant_tolerance_fire_fire_retardant", "plant_tolerance_fire_thick_bark", "flower_orientation_up", "flower_orientation_lateral", "flower_orientation_down", "flower_orientation_mixed", "perianth_colour_white", "perianth_colour_yellow", "perianth_colour_orange", "perianth_colour_red", "perianth_colour_pink", "perianth_colour_purple", "perianth_colour_blue", "perianth_colour_green", "perianth_colour_grey", "perianth_colour_brown", "perianth_colour_black") %>%
paste0(base_url, . )


  to_match_tmp <- triples_with_labels %>% 
    filter(Subject_stripped %in% categorical_allowed) %>%
    filter(property == "label") %>%
    select(Subject_stripped, value)
  
  tmp_categorical2 <- triples_with_labels %>%
    filter(Subject_stripped %in% to_match_tmp$Subject_stripped) %>%
    filter(property == "has broader") %>%
    select(Subject_stripped, trait_name = value) %>%
    left_join(to_match_tmp)
 
  traits_names_tmp <- tmp_categorical2 %>%
    distinct(trait_name)

for (j in seq_along(1:nrow(traits_names_tmp))) {

  writeLines(paste0("\n## ", traits_names_tmp$trait_name[j], "\n\n"))
  
  tmp_cat_select <- tmp_categorical2 %>% 
    filter(trait_name == traits_names_tmp$trait_name[j])
  
  cat_for_trait <- as.data.frame(categorical_allowed) %>% filter(categorical_allowed %in% tmp_cat_select$Subject_stripped)
  
  for (i in seq_along(1:nrow(cat_for_trait))) {
  
  term_to_display <- to_match_tmp$value[match(cat_for_trait$categorical_allowed[i], to_match_tmp$Subject_stripped)]
  
  anchor <- gsub(base_url, "", cat_for_trait$categorical_allowed[i], fixed = TRUE)    
    
  writeLines(paste0("\n### [", term_to_display, "](#", anchor, ")\n\n"))
  
  thistrait <- anchor
      
  out <- 
      create_APD_categorical_values_table(thistrait, triples_with_labels)
  
    out %>% 
      gt() %>% 
      cols_align(align = "left") %>%
      cols_width(
        name ~ px(150),
        description ~px(500)
        ) %>%
      cols_label(
        name = "**Property**",
        description = "**Value**"
        ) %>%
      print()
  }
  
}
  
```

# 5. Glossary

This section of the ontology includes terminology used within trait concept descriptions that are not defined by other resources.

```{r, results='asis', echo=FALSE, message=FALSE, warning=FALSE}
#glossary_terms <- datapasta::vector_paste(glossary$Entity)

base_url <- "https://w3id.org/APD/glossary/"

glossary_terms <- c("glossary_40004", "glossary_40006", "glossary_40014", "glossary_40026", "glossary_40040", "glossary_40044", "glossary_40050", "glossary_40058", "glossary_40060", "glossary_40062", "glossary_40066", "glossary_40068", "glossary_40074", "glossary_40076", "glossary_40080", "glossary_40081", "glossary_40086", "glossary_40088", "glossary_40092", "glossary_40094", "glossary_40106", "glossary_40112", "glossary_40114", "glossary_40124") %>%
paste0(base_url, . )

for (thistrait in glossary_terms) {
  
  to_match_tmp <- triples_with_labels %>% 
    filter(Subject_stripped %in% glossary_terms) %>%
    filter(property == "label") %>%
    select(Subject_stripped, value)
  
  term_to_display <- to_match_tmp$value[match(thistrait, to_match_tmp$Subject_stripped)]
  anchor <- gsub(base_url, "", thistrait, fixed = TRUE)
  
  writeLines(paste0("\n## [", term_to_display, "](#", anchor, ")\n\n"))
  
  out <- 
    create_APD_trait_glossary_table(thistrait, triples_with_labels)

  out %>% 
    gt() %>% 
    cols_align(align = "left") %>%
      cols_width(
        name ~ px(150),
        description ~px(500)
        ) %>%
      cols_label(
        name = "**Property**",
        description = "**Value**"
        ) %>%
    print()
}
```

# 6. Acknowledgements

We are grateful to S Cox, K Levett, M Barlow, C Brady for useful conversations. The AusTraits project received investment (<https://doi.org/10.47486/TD044>, <https://doi.org/10.47486/DP720>) from the Australian Research Data Commons (ARDC). The ARDC is funded by the National Collaborative Research Infrastructure Strategy (NCRIS). 

