---
title: The AusTraits Plant Dictionary (APD)
# The landing page: release metadata, then a filterable table of all 559 trait
# concepts, each row linking to that trait's own page.
#
# It deliberately does NOT carry the whole dictionary. That is full.html, which is
# still published and still archived in each release -- Wenk et al. 2024 (p.12)
# commits to a compiled human-readable version, so it is part of the contract.
# What changed is that looking up one trait no longer means downloading all of it.
#
# No `embed-resources` here: that inlines ~2.4 MB of Quarto's CSS and JS, which is
# 27% of what the old single page weighed.
format:
  html:
    toc: false
include-in-header:
  # Synchronous, so a legacy bookmark redirects before anything renders.
  text: |
    <script>
    (function () {
      var slug = window.location.hash.replace(/^#/, '');
      if (!slug || !/^[A-Za-z][A-Za-z0-9_.-]*$/.test(slug)) return;
      var dir = /^glossary_/.test(slug) ? 'glossary/' : 'traits/';
      window.location.replace(dir + slug + '.html');
    })();
    </script>
---

```{r, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)

source("R/version.R")
source("R/pages.R")
source("R/browse.R")

version <- apd_version()
previous_version <- apd_previous_version(version)

link_version <- function(v) sprintf("<https://w3id.org/APD/release/%s/index.html>", v)

traits <- read_csv("APD_traits.csv", show_col_types = FALSE)
browse <- apd_browse_table(traits)
```

{{< include _release-metadata.qmd >}}

## About

The AusTraits Plant Dictionary (APD) defines the trait concepts used by
[AusTraits](https://austraits.org), a database of Australian plant traits. It covers
`r nrow(browse)` traits spanning plant morphology, nutrient concentrations, physiology,
life history and fire response. Every definition has been reviewed by multiple people
and carries allowable values and ranges, preferred units, keywords, references, and
links to matching traits in other databases.

The APD is a community resource — [suggestions and
corrections](https://github.com/traitecoevo/APD/issues) are welcome, especially
additional references or other databases we should link to.

## Browse traits

Start typing to filter. Every term matches independently, so `leaf nitrogen` finds
traits with both words anywhere in the row. Select a trait to see its full
definition, keywords, references and reviewers.

```{r, results='asis', echo=FALSE}
cat(apd_browse_html(browse))
```

## The rest of the dictionary

| | |
|---|---|
| [Full dictionary](full.html) | Every trait concept, trait group, categorical value and glossary term in one page. This is the compiled human-readable version described in the paper, and the copy archived with each release. |
| [Trait groups](full.html#trait-groups) | The `r sum(!is.na(traits$trait_groupings))` traits organised into broader groupings. |
| [Categorical values](full.html#values-for-categorical-traits) | The allowed values for categorical traits, each defined in its own right. |
| [Glossary](full.html#glossary) | Terms used as keywords that were not available in other published vocabularies. |
| [Using the APD](using_the_APD.html) | Worked examples for searching and extracting the definitions programmatically. |

## Acknowledgements

The APD was developed with support from the Australian Research Data Commons
(ARDC) and the Australian Research Council (ARC). See the
[paper](https://doi.org/10.1038/s41597-024-03368-z) for the full list of
contributors and reviewers.
